\pkg{evtree}: Evolutionary Learning of Globally Optimal Classification and Regression Trees in \proglang{R}

### 2011-09-28 pkg/evtree/inst/doc/evtree.Rnw
added NEWS file, touch-ups to D/M notation in vignette
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@252 255f4249-2263-4417-b1ea-8498779e27a7

\setkeys{Gin}{width=0.5\textwidth}
\begin{figure}[t!]
\centering
<<chess22-plot, fig=TRUE, echo=FALSE, height=4, width=4>>=
plot(X2 ~ X1, data = chess22, xlim = c(0, 2), ylim = c(0, 2), pch = c(1, 4)[Y], col = c("black", "slategray")[Y])
@
\caption{\label{fig:illustration} Class distribution of  the $(X_1, X_2)$-plane. The two classes are indicated by black circles and gray crosses.}
\end{figure}

@@

When compared with the described forward stepwise search, a less greedy approach
is to calculate the effects of the split rules deeper down in the
tree. In this way optimal trees can be found for simple problems. However,
split selection at a given node in Equation~\ref{eq:locally} has complexity  $O(PN)$
(if all $P$~variables are numeric/ordered with $N$ distinct values).
Through a global search up to $D-1$ levels (\ref{eq:globally}) the complexity
increases to $O(P^D N^D)$ \citep{papagelis2001bdt}. One conceivable compromise
between these two extremes is to look ahead $d$~steps with $1 < d < D$
\citep[see e.g.,][]{esmeir2007ald}, also yielding a locally optimal tree but
less constrained than that from a 1-step-ahead search.
Another class of algorithms is given by stochastic optimization methods that,
given an initial tree, seek improved solutions through stochastic changes to
the tree structure. Thus, these algorithms try to explore the full parameter
space $\Theta$ but cannot be guaranteed to find the globally optimal solution
but only an approximation thereof. Besides evolutionary algorithms
\citep{koza1991concept}, \textit{Bayesian CART} \citep{denison1998bca} and
\textit{simulated annealing} \citep{sutton1992ict} were used successfully to
solve difficult classification and regression tree problems.
\cite{koza1991concept} first formulated the concept of using evolutionary
algorithms as a stochastic optimization method to build classification and
regression trees. \cite{papagelis2001bdt} presented a classification tree
algorithm and provided results on several datasets from the UCI machine learning
repository \citep{Frank+Asuncion:2010}. Another method for the construction of
classification and regression trees via evolutionary algorithms was introduced by
\cite{gray2008cta} and \cite{fan2005rta}, respectively.
\cite{cantu2003inducing} used an evolutionary algorithm to induce so-called oblique
classification trees.

>>>>

\setkeys{Gin}{width=0.5\textwidth}
\begin{figure}[t!]
\centering
<<chess22-plot, fig=TRUE, echo=FALSE, height=4, width=4>>=
plot(X2 ~ X1, data = chess22, xlim = c(0, 2), ylim = c(0, 2), pch = c(1, 4)[Y], col = c("black", "slategray")[Y])
@
\caption{\label{fig:illustration} Class distribution of  the $(X_1, X_2)$-plane. The two classes are indicated by black circles and gray crosses.}
\vspace*{-0.3cm} %%FIXME: to avoid \pagebreak
\end{figure}

@@

When compared with the described forward stepwise search, a less greedy approach
is to calculate the effects of the split rules deeper down in the
tree. In this way optimal trees can be found for simple problems. However,
split selection at a given node in Equation~\ref{eq:locally} has complexity  $O(PN)$
(if all $P$~variables are numeric/ordered with $N$ distinct values).
Through a global search up to $D$ levels -- i.e., corresponding to a full binary
tree with $M = 2^D$ terminal nodes -- the complexity
increases to $O(P^D N^D)$ \citep{papagelis2001bdt}. One conceivable compromise
between these two extremes is to look ahead $d$~steps with $1 < d < D$
\citep[see e.g.,][]{esmeir2007ald}, also yielding a locally optimal tree but
less constrained than that from a 1-step-ahead search.

Another class of algorithms is given by stochastic optimization methods that,
given an initial tree, seek improved solutions through stochastic changes to
the tree structure. Thus, these algorithms try to explore the full parameter
space $\Theta$ but cannot be guaranteed to find the globally optimal solution
but only an approximation thereof. Besides evolutionary algorithms
\citep{koza1991concept}, \textit{Bayesian CART} \citep{denison1998bca} and
\textit{simulated annealing} \citep{sutton1992ict} were used successfully to
solve difficult classification and regression tree problems.
\cite{koza1991concept} first formulated the concept of using evolutionary
algorithms as a stochastic optimization method to build classification and
regression trees. \cite{papagelis2001bdt} presented a classification tree
algorithm and provided results on several datasets from the UCI machine learning
repository \citep{Frank+Asuncion:2010}. Another method for the construction of
classification and regression trees via evolutionary algorithms was introduced by
\cite{gray2008cta} and \cite{fan2005rta}, respectively.
\cite{cantu2003inducing} used an evolutionary algorithm to induce so-called oblique
classification trees.
